---
title: "Visualise_staff_Survey"
author: "Marianne van der Vaart"
date: "2024-05-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(ggplot2)
library(tidyverse)
library(hash)
library(cowplot)
library(data.table)
library(plotly)
library(ggpubr)
library(tidyselect)
library(pracma)

genders <- c("Male","Female")
role <- "all"
StaffSurveyTable <- list()

```

This script plots data from the staff survey by year. You will need one excel file per year, where each tab contains the answers for a given gender and role. Example (fake) data is provided in "2021_Staff_Survey_Example.xlsx" and "2023_Staff_Survey_Example.xlsx". Additionally, you will need an excel file which contains the questions you want to plot. The script will look for the question as asked in the first column of this spreadsheet. If the questions change between different years, you must add the alternative question in the second column. An example is provided in "Core_Questions_Example.xlsx". The spreadsheets do not contain real data. 

```{r 2021 & 2023}
years <- c("2021","2023")

for (iYear in 1:length(years))
  {year <- years[iYear]

  # load appropriate file
  filename <- paste0(year, "_Staff_Survey_Example.xlsx")
  
  # loop through genders
  for (iGender in 1:length(genders))
    {gender <- genders[iGender]
    TempTable <- read_xlsx(filename, sheet = paste(gender, role))
    
    TempTable <- TempTable[!is.na(TempTable[,2]),] # remove empty rows
    TempTable <- TempTable[,1:7] # select relevant columns
    TempTable <- mutate(TempTable, year = year) # add year as a variable
    TempTable <- mutate(TempTable, gender = gender) # add gender as a variable
    TempTable <- mutate(TempTable, role = role) # add role as a variable 
    StaffSurveyTable <- rbind(StaffSurveyTable, TempTable)} # store 
  }

StaffSurveyTable <- melt(setDT(StaffSurveyTable), id.vars = c("Theme","Question","Population","Percentage Score", "year","gender","role"), variable.name = "Answer") # convert to long format
StaffSurveyTable <- rename(StaffSurveyTable, Percentage = value) # rename the percentage variable
 
```


```{r Bind}

# Organise for plotting
StaffSurveyTable$Percentage <- as.numeric(StaffSurveyTable$Percentage) * 100  
StaffSurveyTable$Answer <- factor(StaffSurveyTable$Answer, levels = c("Negative","Neutral","Positive"))
StaffSurveyTable$gender <- factor(StaffSurveyTable$gender, levels = genders)
```


```{r All Qs}
years <- c("2021","2023")
cols <- c("#80b1d3","#fb8072")
p <- list()

# Load in the relevant questions
questions <- as.data.frame(read_xlsx("Core_questions_Example.xlsx"))

# Specify which percentages to plot
plottextfor <- c("Positive","Negative")

# Loop through questions
for (iQuestion in 1:length(questions$Questions))
{question <- questions[iQuestion,1:2] # select question
  data <- StaffSurveyTable

  # Questions may change year by year, so find the 2023 question or its alternative
  data1 = data[grepl(question$Questions2023, data$Question, fixed = TRUE)] # grab relevant question from data frame (2023 questions)
  data2 = data[grepl(question$Alternative2021, data$Question, fixed = TRUE)] # grab relevant question from data frame (alternative - 2021 questions)
  data = rbind(data1, data2)
  
  # Loop through years
  for (iYear in 1:length(years))
    {
    x1 <- data[data$year == years[iYear],] # select relevant data
    x1 <- x1[!is.na(x1$Percentage)] # remove missing data
    x1 <- group_by(x1, gender) # group by gender
    x1 <- mutate(x1, label_y_pos = cumsum(Percentage)) # sum the percentages to find where to plot them
    x1 <- mutate(x1, label_y = round(Percentage,0)) # round them
    x1$label_y <- ifelse(x1$Answer %in% plottextfor, x1$label_y, " ")
    x1$gender_n <- paste0(x1$gender, ", n = ", x1$Population)
    
    # create the plot
  p[[iYear]] <- ggplot(x1, aes(x = gender, y = Percentage)) +geom_col(aes(fill = interaction(gender,Answer)), width = 0.7)+ geom_text(aes(y = label_y_pos, label = label_y,hjust = 1.5), colour = "white") + coord_flip() + ggtitle(paste(years[iYear], x1$Question[1]))+ theme(plot.title = element_text(size = 9.5, face = "bold"), legend.title = element_blank())+ scale_fill_manual(values=c("#D7E3EC","#FFE4E1","#B1C7D8","#FFC1BA","#80b1d3","#fb8072"))+theme(legend.position = "bottom")+labs(x = "Gender", colour = "Answer")+ theme(axis.title = element_text(size = 9), axis.text.y = element_text(size = 8), legend.text = element_text(size = 8))
  }
  
# Remove the labels from the top plot
p[[1]] <- p[[1]] + theme(axis.title.x = element_blank(),axis.text.x = element_blank(),axis.ticks.x = element_blank(), legend.position = "none")

# Combine the plots
pfull <- plot_grid(p[[1]],p[[2]], nrow=2,  rel_heights = c(5,8))

# Save 
print(pfull)
save_plot(paste0(question$Questions2023,"_",role,".png"), pfull, base_width = 9, base_height=4, bg='white')

}

```
          


